# pre-class


Make sure you commit this often with meaningful messages. 

```{r}
library(dplyr)
```




### Question 1:

Using a loop, print the integers from 1 to 50. 
```{r}

for(i in 1:50){ #using i itself as the value to print
  print(i)
}


```
### Question 2:

A.  Using a loop, add all the integers between 0 and 1000.
```{r}
sum <- 0 #Set sum = 0 to start
for(i in 1:1000){
  sum <- sum+i #Add sum to to i as we let i go from one to one thousand
}
sum #Print for confirmation
```




B. Now, add all the EVEN integers between 0 and 1000 (hint: use seq())
```{r}
sum2 <- 0 #Set sum2 = 0 to start
for(i in seq(from =0, to = 1000, by = 2)){#This will go 0, 2, 4, 6... to 1000.
  sum2 <- sum2+i
}
sum2 #Print for confirmation
```
C. Now, repeat A and B WITHOUT using a loop.
```{r}
#Set sum2,sum3 = 0 to start
sum3 <- 0
sum4 <- 0

sum3 <- sum3+seq(from =0, to = 1000, by = 1)#This will go 0, 1, 2, 3... to 1000.
#sum3 is now a list of the numbers we want to sum so...
sum(sum3)

sum4 <- sum4+seq(from =0, to = 1000, by = 2)#This will go 0, 2, 4, 6... to 1000.
#sum4 is now a list of the numbers we want to sum so...
sum(sum4)
```

### Question 3:

Here is a dataframe of survey data containing 5 questions :

```{r, eval=FALSE}
survey <- data.frame( #All of this code is given
                     "participant" = c(1, 2, 3, 4, 5, 6),
                     "q1" = c(5, 3, 2, 7, 11, 0),
                     "q2" = c(4, 2, 2, 5, -10, 99),
                     "q3" = c(-4, -3, 4, 2, 9, 10),
                     "q4" = c(-30, 5, 2, 23, 4, 2),
                     "q5" = c(88, 4, -20, 2, 4, 2)
                     )
```
The response to each question should be an integer between 1 and 5. Obviously, we have some bad values in the dataframe. The goal of this problem is to fix them.

A. Using a loop, create a new dataframe called survey.clean where all the invalid values (those that are not integers between 1 and 5) are set to NA.
```{r}
survey.clean<- survey #Define our new data frame to edit

  for (i in 1:6){ 
  for (j in 2:6){ # Loop through the columns and rows skipping participant #6
    
    if(survey.clean[i,j] < 1 | survey.clean[i,j] > 5){ #Loop through each element using i and j themselves
      survey.clean[i,j]=NA #Set the searched for values to NA
      
    }
  }}

survey.clean #print to check
```
B. Now, again using a loop, add a new column to the dataframe called “invalid.answers” that indicates, for each participant, how many bad answers they gave.
```{r}

#We're using survey.clean here so rerun the code for part A eachtime this is run or it'll overwrite your data

NASUM <- 0 #We'll use this to keep track of the bad answers for each row

  for (i in 1:6){
  for (j in 1:6){ # Loop through the columns and rows 
    
    if(is.na(survey.clean[i,j])){ # So for each NA we find...
      NASUM <- NASUM +1 #We add 1 to NASUM for this row then...
    }
  }
   survey.clean[i,7] = NASUM #We set our new row equal to NASUM
   NASUM <- 0 #We reset NASUM to 0 so we can go through another row
  }

names(survey.clean)[7]<-"invalid.answers" #We need to name this row as specified in the question prompt

survey.clean #print to check
```