# pre-class


Make sure you commit this often with meaningful messages. 



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```








Standardizing a variable means subtracting the mean, and then dividing by the standard deviation. Let’s use a loop to standardize the numeric columns in the [Western Collaborative Group Study](https://clinicaltrials.gov/ct2/show/NCT00005174). This study began in 1960 with 3154 men ages 39-59, who were employed in one of 11 California based companies. They were followed until 1969 during this time, 257 of these men developed coronary heart disease (CHD). You can read this data in with the code below. You can access this dataset with the following code:

```{R}

suppressMessages(library(foreign))
wcgs <- read.dta("https://drive.google.com/uc?export=download&id=0B8CsRLdwqzbzYWxfN3ExQllBQkU")
```

The data has the following variables:



WCGS has the following variables:

-----------------------------------------------------------
Name    Description
------- -------------------------------------------
id      Subject identification number

age     Age in years

height  Height in inches

weight  Weight in lbs.

sbp     Systolic blood pressure in mm 

dbp     Diastolic blood pressure in mm Hg

chol    Fasting serum cholesterol in mm 

behpat  Behavior

  1       A1

  2       A2

  3       B3

  4       B4

ncigs   Cigarettes per day

dibpat  Behavior

1       type A

2       type B

chd69   Coronary heart disease

1       Yes

0       no

typechd Type of CHD

1       myocardial infarction or death

2       silent myocardial infarction

3       angina perctoris

time169 Time of CHD event or end of follow-up

arcus   Arcus senilis

0       absent

1       present

bmi     Body Mass Index
-----------------------------------------------------------




### Question 1: Standardize Function

A. Create a function called standardize.me() that takes a numeric vector as an argument, and returns the standardized version of the vector. 

```{R}
standardize.me <-function(Vector){

StandardizedVector <- (Vector - mean(Vector,na.rm=TRUE))/(sd(Vector,na.rm=TRUE)) #Standardizing the vector

return(StandardizedVector)} #Don't forget to return this
```

}
B. Assign all the numeric columns of the original WCGS dataset to a new dataset called WCGS.new.
```{r}
WCGS.new <-wcgs
```


C. Using a loop and your new function, standardize all the variables WCGS.new dataset.
```{r}
for(i in 1:16){
  if(is.numeric(WCGS.new[[i]])){
    WCGS.new[[i]]<-standardize.me(WCGS.new[[i]])
  }
}

```

D. What should the mean and standard deviation of all your new standardized variables be? Test your prediction by running a loop

**These should all be 0 and 1 respectively, let's see if we can get this**

```{r}
for(i in 1:16){
  if(is.numeric(WCGS.new[[i]])){
    mean <-mean(WCGS.new[[i]],na.rm=TRUE)
    sd <-sd(WCGS.new[[i]],na.rm=TRUE)
    name<- colnames(wcgs[i])
    print(paste("Mean for ",name,"is ",mean))
    print(paste("Standard deviation for ",name,"is ",sd))
  }
}
```


### Question 2: Looping to Calculate

A. Using a loop, calculate the mean weight of the subjects separated by the type of CHD they have.

```{r}
#I'm going to define two empty vectors, and as I loop through wcgs fill the "yes" vector with the weights where chd69 is yes and fill the "no" vector with the weights where chd69 is no

noVector<-c() 
noCount<-0 #Counter is needed to keep track of the elements in the vecttor

yesVector<-c()
yesCount<-0 #Counter is needed to keep track of the elements in the vecttor

for(i in 1:nrow(wcgs)){
if(wcgs$chd69[i]=="no"){
  noVector[noCount] <- wcgs$weight[i] #Add the ith element of wcgs$weight to noVector as the 'noCount'th element
  noCount<-noCount+1 #add one to our count variable so we don't overwrite our values
}else if (wcgs$chd69[i]=="yes"){
  yesVector[yesCount] <- wcgs$weight[i] #Add the ith element of wcgs$weight to yesVector as the 'yesCount'th element
  yesCount<-yesCount+1 #add one to our count variable so we don't overwrite our values
}
}

meantable<-c(mean(noVector),mean(yesVector))
names(meantable)<-c("Mean of Yes","Mean of No")
meantable
```
B. Now do the same thing, but now don’t use a loop
```{r}
#library(dplyr)

wcgs %>%
  group_by(chd69) %>% #Groups our yes and no variables together and then...
  summarise(mean(weight))#We take their mean!
```

